---
format:
  gfm:
    default-image-extension: ""
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pidigits

<!-- badges: start -->
<!-- badges: end -->

**pidigits** provides an arbitrary-precision helper function, `pi_digits()`, for computing decimal digits of $\pi$ in R.

Under the hood, calculations are powered by GNU MPFR via the [Rmpfr](https://cran.r-project.org/package=Rmpfr) library. Results are automatically cached with [memoise](https://cran.r-project.org/package=memoise), ensuring that repeated queries for the same precision are instantaneous.

## Features

- **Arbitrary Precision**: Compute dozens, hundreds, or thousands of decimal digits of $\pi$.
- **Flexible Return Types**:
  - `"char"` (default): Returns a character string (`"3.14159..."`), ensuring exact digits are preserved without floating-point truncation.
  - `"numeric"`: Converts the result to standard R 64-bit double precision.
  - `"mpfr"`: Returns an `Rmpfr` arbitrary-precision object for downstream mathematical operations.
- **Automatic Memoisation**: Results are cached in memory so subsequent calls for the same $n$ take zero computation time.

## Installation

You can install the development version of pidigits from [GitHub](https://github.com/jkylearmstrong/pidigits) with:

```r
# install.packages("pak")
pak::pak("jkylearmstrong/pidigits")
```

## Quick Start

```{r}
#| label: example
library(pidigits)

# First 10 decimal digits (as character string)
pi_digits(10)

# 50 decimal digits
pi_digits(50)

# Return as standard R double numeric
pi_digits(20, type = "numeric")

# Return as an Rmpfr arbitrary precision object
pi_digits(50, type = "mpfr")
```

## Memoisation in Action

Because `pi_digits()` is memoised, repeated calls for the same digit length return instantly from the cache:

```{r}
#| label: memoise-demo
# Initial computation (computes and caches)
system.time(pi_digits(2000))

# Second call (instant cache hit)
system.time(pi_digits(2000))
```

## Development

`README.md` is generated from `README.qmd`. To regenerate `README.md`, run:

```r
devtools::build_readme()
```
or render with Quarto:

```bash
quarto render README.qmd
```
